\Huge Literature Search Made Easy

### 2013-12-14 doc/article_1.tex
Minor changes
Final changes

% The first character should be within \initial{}
\initial{M}\textbf{any of us have been faced with this situation: we found a certain field of research intriguing, went to an online archive to search for related papers, and the results were just less than satisfactory: they are too unstructured to obtain a big picture, such as how this field evolves over time, what sub-fields it is composed of. Instead of relying on human instinct for such insights, our project provides an alternative: using machine learning algorithms to peruse thru online archives, generate the information of interest, and present them in a human friendly way. In this project, we used hierarchical latent Dirichlet allocation (hLDA) to find the topics and their hierarchy and dynamic topic model (DTM) to find the time evolution of topics. In addition to using these two models, we also made an important extension to them, i.e. constructing the corpus with "key phrases" instead of "keywords" . Furthermore, we devised an algorithm to remove the duplication in word and phrase count, which enabled us to generate a corpus composed of both keywords and "key phrases". This approach allowed us to greatly improve the interpretability of the result since a large number of concepts in science and technology are formulated as phrases.\\
In our experiments with 14185 condensed matter physics papers from arXiv, the above methods effectively found the trend and hierarchy of research topics and revealed some very interesting insights. In our report, the "Trends in Research Topics" section demonstrates the analysis of topic trends, advantage of using "key phrases" versus keywords and a caveat in word and phrase count when constructing the corpus. The "Topic Hierarchy" section presents the generation of a hierarchy of topics and the solution to the caveat aforementioned.}
%----------------------------------------------------------------------------------------
%	ARTICLE CONTENTS
%----------------------------------------------------------------------------------------

@@

\subsection*{Phrase Generation}
While doing data pre-processing, we observed that new concepts in science and technology are usually expressed in the form of phrases, such as "computer science", "machine learning" and "topic modeling", which indicates a topic might be better characterized by its "key phrases" instead of keywords. Furthermore, phrases also offer better interpretability of results, e.g. a topic with keywords {"dimension", "principal", "analysis"} may greatly perplex us, while one with key phrases {"principal component analysis", "dimension reduction"} is quite informative.\newline
We generated phrases before the second step in the previous section by:
\begin{enumerate}
  \item Separate words into blocks by punctuation marks and stop words.
  \item Within each block, generate phrases made up of consecutive words while restricting the maximum length of phrases to 5 words since usually phrases longer than 5 words are represented by their acronyms.
\end{enumerate}
\subsection*{Duplicate Counts Removal}
Using the phrase generation method described above can lead to duplicate counts in the resulting corpus. For example, phrase "Stanford University" will generate "Stanford" once, "University" once and "Stanford University" once, while these two words should only be counted as a phrase once. To eliminate this inaccuracy, we implement a two-phase algorithm. The first phase is the same as the approach described above: extract words and enumerate phrases in each document, then filter the words and phrases whose occurrences are below the minimum threshold. In the second phase, documents are scanned through again. For each word in a document, we check whether it can be combined with its previous or next word to generate a phrase appearing in the filtered phrase dictionary. If not, this word is counted as one individual occurrence; otherwise it is skipped. This method basically recounts all terms and checks if their appearance contribute to the generation of a phrase. Only those terms not as part of a phrase is counted. By using this procedure, we were able to produce a corpus which nicely combines both phrases and words.

@@

In our DTM experiment, the number of topics is set to 10 a priori. Results are summarized in Figure ~\ref{fig: DTM_table_unweight} and Figure ~\ref{fig: PhaseTrans_wp}. In Figure ~\ref{fig: DTM_table_unweight} we demonstrate 5 out of the 10 topics for purpose of figure size control, the 3 tables respectively are topics and keywords generated from only words, words plus phrases and only phrases. In these tables, the topics are written in bold and keywords in regular font style. We see the in all cases, the DTM algorithm very nicely identifies major research topics. However, the classification with phrases only clearly gives much better interpretability. Note that the first two tables look quite similar, in fact, in the second table, even though the corpus contains both words and phrases, we can hardly see any phrases among the important keywords. This is quite intuitive since in the corpus, there are 486712 words, 4645 of them unique and 112448 phrases with 11250 of them unique. So on average, a word appears ~100 times in the corpus while a phrase appears only ~10 times, resulting in a term matrix dominated by words. This is what has motivated to remove the duplicated counts in words.

@@

From Figure ~\ref{fig: PhaseTrans_wp}, despite the frequencies of keywords do rise and fall, there is hardly any obvious "trend". This is not unexpected since it's natural that a paper on "phase transition" should always involve keywords such as "phase transition" and "phase diagram" regardless of when it is published.\newline
One very informative plot we can make is the evolution of topic popularity. In Figure ~\ref{fig: TopPop_p}, we show a plot of frequencies for 4 topics over time.

@@

\begin{inparaenum}[\itshape a\upshape)]

@@

\end{inparaenum}

\subsubsection*{Topic Hierarchy}
 We produced four depth 4 trees generated by corpora composed of
 \begin{inparaenum}[\itshape a\upshape)]
 \item  words alone
 \item phrases alone
 \item words plus phrases without duplicate count removal
 \item words plus phrases with duplicate count removal
 \end{inparaenum}
. As expected, results from the first and third corpus composition look almost identical, as in the DTM case, so in the report we show only one of them. In Figure ~\ref{fig: Tree_words}, ~\ref{fig: Tree_phrase} and 9, the other three trees are presented, where each node represents a topic. Since these trees have a total of 4 levels and tens of leaves, we include in the plots only 5 or 6 topics on the second level, and expand further one out of those 5 or 6. The remaining topics are represented by $\ldots \ldots$. 
\begin{figure}[!ht]
	\centerline{\includegraphics[scale = 1]{tree_words.eps}}
	\caption{Hierarchical Classification of a Corpus Composed of Words alone}
	 \label{fig: Tree_words}
\end{figure}
\begin{figure}[!ht]
	\centerline{\includegraphics[scale = 0.57]{tree_phrase.eps}}
	\caption{Hierarchical Classification of a Corpus Composed of Phrases alone}
	 \label{fig: Tree_phrase}
\end{figure}
\begin{figure}[!ht]
	\centerline{\includegraphics[scale = 0.57]{tree_words_phrase.eps}}
	\caption{Hierarchical Classification of a Corpus Composed of Words plus Phrases with Removal of Duplication}
	 \label{fig: Tree_phrase}
\end{figure}
All of these three plots look similar on the second level, and the classifications very accurately capture the sub fields in condensed matter physics. This is quite surprising especially considering that hLDA a completely unsupervised learning algorithm.\\
Here again we see that using phrases results in much better interpretability, e.g. in the second tree, phrases in every rectangle look like something meaningful (and indeed they are) even to those who are not experts in physics, while in the first tree the topic with keywords "random, dirac, qcd, eigenvalue" is very bewildering, hence not labeled with any topic. However, there is a draw back of only including phrases in the corpus: words such as "superconductor" and "antiferromaget", though being key concepts in condensed matter physics cannot be used in the classification because they are single words. So in the third tree, we made use of a combination of words and phrases. After removing the duplicated counts in words, the keywords generated by the hLDA algorithm have a well proportioned mixture of words and phrases instead of solely dominated by words. In addition, the topic "superconductivity" is now chacterized by keyword "superconductor" instead of the phrase "superconducting state". Combining words and phrases and implementing duplication removal allows the algorithm to choose the most important expressions from both of two sets instead of one, resulting in better performance and interpretability. Hence we think this is the best way to do topic modeling.

\section*{Conclusion}
In this work, we proposed a phrase-based extension to topic modeling which improves the clustering performance and enables effective interpretation of the document context. Phrases contain important information on the correlation between documents and can help human to understand since many words appear in combinations. The two-phase phrase extraction mechanism first enumerates word combinations and then filters out significant phrases based on cross document validation. This approach implicitly selects the key phrases in the corpus since it leverages the information hidden in multiple documents. In the evaluation, we compared three scenarios, i.e. word-only, phrase-only, and word+phrase. In both hLDA and DTM modeling, we can clearly see that with phrase information the result is much more accurate and easy to interpret.

>>>>

% The first character should be within \initial{}
\initial{M}\textbf{any of us have been faced with this situation: we found a certain field of research intriguing, went to an online archive to search for related papers, and the results were just less than satisfactory: they are too unstructured to obtain a big picture, such as how this field evolves over time, what sub-fields it is composed of. Instead of relying on human instinct for such insights, our project provides an alternative: using machine learning algorithms to peruse thru online archives, generate the information of interest, and present them in a human friendly way.\\
In this project, we used hierarchical latent Dirichlet allocation (hLDA) to find the topics and their hierarchy and dynamic topic model (DTM) to find the time evolution of topics. In addition to using these two models, we also made an important extension to them - constructing the corpus with {\it key phrases} instead of {\it keywords}. Furthermore, we devised an algorithm to remove the duplication in word and phrase count, which enabled us to generate a corpus composed of both keywords and key phrases. This approach allowed us to greatly improve the interpretability of the result since a large number of concepts in science and technology are formulated as phrases.\\
In our experiments with 14185 condensed matter physics papers from arXiv, the above methods effectively found the trend and hierarchy of research topics and revealed some very interesting insights. In our report, the Trends in Research Topics section demonstrates the analysis of topic trends, advantage of using key phrases versus keywords and a caveat in word and phrase count when constructing the corpus. The Topic Hierarchy section presents the generation of a hierarchy of topics and the solution to the caveat aforementioned.}
%----------------------------------------------------------------------------------------
%	ARTICLE CONTENTS
%----------------------------------------------------------------------------------------

@@

\subsection*{Phrase Generation}
While doing data pre-processing, we observed that new concepts in science and technology are usually expressed in the form of phrases, such as "computer science", "machine learning" and "topic modeling", which indicates a topic might be better characterized by its "key phrases" instead of keywords. Furthermore, phrases also offer better interpretability of results, e.g. a topic with keywords {"dimension", "principal", "analysis"} may greatly perplex us, while one with key phrases {"principal component analysis", "dimension reduction"} is quite informative.\newline \newline
We generated phrases before the second step in the previous section by:
\begin{enumerate}
  \item Separate words into blocks by punctuation marks and stop words.
  \item Within each block, generate phrases made up of consecutive words while restricting the maximum length of phrases to 5 words since usually phrases longer than 5 words are represented by their acronyms.
\end{enumerate}
\subsection*{Duplicate Counts Removal}
Using the phrase generation method described above can lead to duplicate counts in the resulting corpus. For example, phrase "Stanford University" will generate "Stanford" once, "University" once and "Stanford University" once, while these two words should only be counted as a phrase once. \newline \newline
To eliminate this inaccuracy, we implement a two-phase algorithm. The first phase is the same as the approach described above: extract words and enumerate phrases in each document, then filter the words and phrases whose occurrences are below the minimum threshold. In the second phase, documents are scanned through again. For each word in a document, we check whether it can be combined with its previous or next word to generate a phrase appearing in the filtered phrase dictionary. If not, this word is counted as one individual occurrence; otherwise it is skipped. This method basically recounts all terms and checks if their appearance contribute to the generation of a phrase. Only those terms not as part of a phrase is counted. By using this procedure, we were able to produce a corpus which nicely combines both phrases and words.

@@

In our DTM experiment, the number of topics is set to 10 a priori. Results are summarized in Figure ~\ref{fig: DTM_table_unweight} and Figure ~\ref{fig: PhaseTrans_wp}.\newline \newline
In Figure ~\ref{fig: DTM_table_unweight} we demonstrate 5 out of the 10 topics for purpose of figure size control, the 3 tables respectively are topics and keywords generated from only words, words plus phrases and only phrases. In these tables, the topics are written in bold and keywords in regular font style. We see the in all cases, the DTM algorithm very nicely identifies major research topics. However, the classification with phrases only clearly gives much better interpretability. Note that the first two tables look quite similar, in fact, in the second table, even though the corpus contains both words and phrases, we can hardly see any phrases among the important keywords. This is quite intuitive since in the corpus, there are 486712 words, 4645 of them unique and 112448 phrases with 11250 of them unique. So on average, a word appears ~100 times in the corpus while a phrase appears only ~10 times, resulting in a term matrix dominated by words. This is what has motivated to remove the duplicated counts in words.\newline \newline

@@

From Figure ~\ref{fig: PhaseTrans_wp}, despite the frequencies of keywords do rise and fall, there is hardly any obvious "trend". This is not unexpected since it's natural that a paper on "phase transition" should always involve keywords such as "phase transition" and "phase diagram" regardless of when it is published.\newpage \noindent
One very informative plot we can make is the evolution of topic popularity. In Figure ~\ref{fig: TopPop_p}, we show a plot of frequencies for 4 topics over time.\par
\begin{figure}[h!]

@@

\begin{enumerate}[\itshape a\upshape)]

@@

\end{enumerate}

\subsubsection*{Topic Hierarchy}
 We produced four depth 4 trees generated by corpora composed of
 \begin{inparaenum}[\itshape a\upshape)]
 \item  words alone
 \item phrases alone
 \item words plus phrases without duplicate count removal
 \item words plus phrases with duplicate count removal
 \end{inparaenum}.\newline \newline
As expected, results from the first and third corpus composition look almost identical, as in the DTM case, so in the report we show only one of them. In Figure ~\ref{fig: Tree_words}, ~\ref{fig: Tree_phrase} and 9, the other three trees are presented, where each node represents a topic. Since these trees have a total of 4 levels and tens of leaves, we include in the plots only 5 or 6 topics on the second level, and expand further one out of those 5 or 6. The remaining topics are represented by "$\ldots \ldots$". 
\begin{figure}[!ht]
	\centerline{\includegraphics[scale = 0.75]{tree_words.eps}}
	\caption{Hierarchical Classification of a Corpus Composed of Words alone}
	 \label{fig: Tree_words}
\end{figure}
\newpage
\begin{figure}[!ht]
	\centerline{\includegraphics[scale = 0.6]{tree_phrase.eps}}
	\caption{Hierarchical Classification of a Corpus Composed of Phrases alone}
	 \label{fig: Tree_phrase}
\end{figure}
\begin{figure}[!ht]
	\centerline{\includegraphics[scale = 0.6]{tree_words_phrase.eps}}
	\caption{Hierarchical Classification of a Corpus Composed of Words plus Phrases with Removal of Duplication}
	 \label{fig: Tree_phrase}
\end{figure}
All of these three plots look similar on the second level, and the classifications very accurately capture the sub fields in condensed matter physics. This is quite surprising especially considering that hLDA a completely unsupervised learning algorithm.\newline \newline
Here again we see that using phrases results in much better interpretability, e.g. in the second tree, phrases in every rectangle look like something meaningful (and indeed they are) even to those who are not experts in physics, while in the first tree the topic with keywords "random, dirac, qcd, eigenvalue" is very bewildering, hence not labeled with any topic. However, there is a draw back of only including phrases in the corpus: words such as "superconductor" and "antiferromaget", though being key concepts in condensed matter physics cannot be used in the classification because they are single words. So in the third tree, we made use of a combination of words and phrases. After removing the duplicated counts in words, the keywords generated by the hLDA algorithm have a well proportioned mixture of words and phrases instead of solely dominated by words. In addition, the topic "superconductivity" is now chacterized by keyword "superconductor" instead of the phrase "superconducting state". Combining words and phrases and implementing duplication removal allows the algorithm to choose the most important expressions from both of two sets instead of one, resulting in better performance and interpretability. Hence we think this is the best way to do topic modeling.

\section*{Conclusion}
In this work, we proposed a phrase-based extension to topic modeling which improves the clustering performance and enables effective interpretation of the document context. Phrases contain important information on the correlation between documents and can help human understand since many words do appear in combinations. The two-phase phrase extraction mechanism first enumerates word combinations and then filters out significant phrases based on cross document validation. This approach implicitly selects the key phrases in the corpus since it leverages the information hidden in multiple documents. In the evaluation, we compared three scenarios, i.e. word-only, phrase-only, and word+phrase. In both hLDA and DTM modeling, we can clearly see that with phrase information the result is much more accurate and easy to interpret.
